tl20-05_berghei_pir_DESEQ

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Less adjustment DESeq and Label height change to fit

Individual _pir_ genes which we use as examples of being highly upregulated in individual stages.

Having looked into independent filter cut-offs, no genes are removed even by ludicrous 'alpha' values, so I assume that our dataset contains few low expressed genes (remember that we filtered low counts out earlier) or outliars of concern.

```{r}
ordered_stages <- ordering(unique(coldata$stage), order_vector = r_order_vector)
comparisons <- c(
  lapply(seq_along(ordered_stages),
         function(i){ if(i < length(ordered_stages)){ c(ordered_stages[i+1],ordered_stages[i]) } }
  ),
  list(
    c('Liv.24h','Liv.post48h'),
    c('Gam.Male','Asex.Schiz'),
    c('Sporoz','Ook'),
    c('Gam.Male','Asex.Mixed'),
    c('Gam.Male','Liv.Meroz')
  )
)
comparisons <- comparisons[-which(sapply(comparisons, is.null))]
#Remove some comparisons
comparisons <- comparisons[!(grepl(comparisons, pattern = 'Asex\\.Mixed') & grepl(comparisons, pattern = 'Asex\\.Ring'))]

shrunk.list <- lapply(comparisons, 
                      function(comparison){
                        res <- results(
                          dds.deseq,
                          contrast = c("stage",comparison[1],comparison[2])
                        )
                        shrunk <- lfcShrink(dds.deseq, 
                                            res = res,
                                            type = "ashr") %>% 
                          as.data.frame %>% 
                          rownames_to_column(var = 'Geneid') %>% 
                          mutate(comp = paste0(comparison[1],'_',comparison[2])) %>% 
                          filter(Geneid %in% pir_id,
                                 !is.na(padj))
                        return(shrunk)
                      })
names(shrunk.list) <- unlist(lapply(comparisons, 
                                    function(comparison) paste0(comparison[1],'_',comparison[2])))
sig_pirs <- bind_rows(shrunk.list) %>% 
  mutate(fdr = p.adjust(pvalue),
         subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
  arrange(subfam, Geneid) %>% 
  filter(fdr < 0.05)
sig_pirs_order <- unique(sig_pirs$Geneid)
sig_pirs_cast <- dcast(sig_pirs,
                  comp ~ Geneid,
                  value.var = 'log2FoldChange') %>% 
  column_to_rownames(var = 'comp')
sig_pirs_cast <- sig_pirs_cast[ordering(rownames(sig_pirs_cast),
                              r_order_vector),
                     sig_pirs_order]

sig_pirs %>% filter(Geneid %in% 'PBANKA_0317181')
sig_pirs %>% filter(Geneid %in% 'PBANKA_0600031')

png(filename = "plots/thesis_chapter3_figure_statsCompHeatmap_inidvidualGenes.png",
    width = 7,
    height = 5,
    units = "in",
    res = 330)
Heatmap(sig_pirs_cast,
        name = 'logFoldChange',
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        # rect_gp = gpar(col = "white", lwd = 1),
        border_gp = gpar(col = "black", lty = 1),
        na_col = 'white',
        show_column_names = FALSE,
        column_split = gsub(pir_clade$clade[match(colnames(sig_pirs_cast), pir_clade$Geneid)],
                            pattern = 'ancestral',
                            replacement = 'a')
)
dev.off()
```

>>>>

Having looked into independent filter cut-offs, no genes are removed even by ludicrous 'alpha' values, so I assume that our dataset contains few low expressed genes (remember that we filtered low counts out earlier) or outliers of concern.

```{r comparisons_ordering}
ordered_stages <- ordering(unique(coldata$stage), order_vector = r_order_vector)
comparisons <- c(
  lapply(seq_along(ordered_stages),
         function(i){ if(i < length(ordered_stages)){ c(ordered_stages[i+1],ordered_stages[i]) } }
  ),
  list(
    c('Liv.post48h','Liv.24h'),
    c('Gam.Male','Asex.Schiz'),
    c('Sporoz','Ook'),
    c('Gam.Male','Asex.Mixed'),
    c('Gam.Male','Liv.Meroz')
  )
)
comparisons <- comparisons[-which(sapply(comparisons, is.null))]
#Remove some comparisons
comparisons <- comparisons[!(grepl(comparisons, pattern = 'Asex\\.Mixed') & grepl(comparisons, pattern = 'Asex\\.Ring'))]
```

```{r lfcShrink}
shrunk.list <- lapply(comparisons, 
                      function(comparison){
                        res <- results(
                          dds.deseq,
                          contrast = c("stage",comparison[1],comparison[2])
                        )
                        shrunk <- lfcShrink(dds.deseq, 
                                            res = res,
                                            type = "ashr") %>% 
                          as.data.frame %>% 
                          rownames_to_column(var = 'Geneid') %>% 
                          mutate(comp = paste0(comparison[1],'_',comparison[2])) %>% 
                          filter(Geneid %in% pir_id,
                                 !is.na(padj))
                        return(shrunk)
                      })
names(shrunk.list) <- unlist(lapply(comparisons, 
                                    function(comparison) paste0(comparison[1],'_',comparison[2])))
sig_pirs <- bind_rows(shrunk.list) %>% 
  mutate(subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
  arrange(subfam, Geneid) %>% 
  filter(padj < 0.05)
sig_pirs_order <- unique(sig_pirs$Geneid)
sig_pirs_cast <- dcast(sig_pirs,
                       comp ~ Geneid,
                       value.var = 'log2FoldChange') %>% 
  column_to_rownames(var = 'comp')
sig_pirs_cast <- sig_pirs_cast[ordering(rownames(sig_pirs_cast),
                                        c(r_order_vector,
                                          'Gam.Male_Asex.Mixed',
                                          'Gam.Male_Liv.Meroz')),
                               sig_pirs_order]

sig_pirs %>% filter(Geneid %in% 'PBANKA_0317181')
sig_pirs %>% filter(Geneid %in% 'PBANKA_0600031')
```

```{r heatmap}
#Add a new-line to the S6 label so that it doesn't overlap with others
pir_column_split <- gsub(pir_clade$clade[match(colnames(sig_pirs_cast), pir_clade$Geneid)],
                         pattern = 'ancestral',
                         replacement = 'a')
pir_column_split <- ifelse(pir_column_split %in% c('S6', 'L2'), 
                           paste0(pir_column_split, '\n'),
                           pir_column_split)

png(filename = "plots/thesis_chapter3_figure_statsCompHeatmap_inidvidualGenes.png",
    width = 7,
    height = 5,
    units = "in",
    res = 330)
Heatmap(sig_pirs_cast,
        name = 'logFoldChange',
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        # rect_gp = gpar(col = "white", lwd = 1),
        border_gp = gpar(col = "black", lty = 1),
        column_title_gp = gpar(fontsize = 12),
        na_col = 'white',
        show_column_names = FALSE,
        column_split = pir_column_split,
        heatmap_height = unit(10, "cm"),
        #Shrink heatmap a little so that the S6 label is visible
        column_gap = unit(2, 'mm')
)
dev.off()
```
